Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-05 paper/main.tex
shorten the abstract

% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
% \institute{
%    Basque Center for Applied Mathematics\\
%    \email{eirurozki@bcamath.org}
%    \and
%    University of Málaga, Málaga, Spain\\
%    \email{manuel.lopez-ibanez@uma.es}
%  }
%    \renewcommand{\shortauthors}{B. Trovato et al.}
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Expensive black-box combinatorial optimization problems arise in practice
  when the objective function is evaluated by means of a simulator or a
  real-world experiment. Since each fitness evaluation is expensive in terms of
  time or resources, only a limited number of evaluations is possible,
  typically several orders of magnitude smaller than in non-expensive
  problems. In this scenario, classical optimization methods such as
  mixed-integer programming and local search are not useful.  In the continuous
  case, Bayesian optimization, in particular using Gaussian processes, has
  proven very effective under these conditions. Much less research is available
  in the combinatorial case. In this paper, we propose and analyze UMM, an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and unbalanced rank aggregation (uBorda).

  Experimental results on black-box versions of LOP and PFSP show that UMM
  outperforms the solutions obtained by CEGO, a Bayesian optimization algorithm
  for combinatorial optimization. Nevertheless, a slight modification to CEGO,
  based on the different interpretations for rankings and orderings,
  significantly improves its performance, thus producing solutions that are
  slightly better than those of UMM and dramatically better than the original
  version.  As another benefit, the computational complexity of UMM increases
  linearly with both the number of function evaluations and the permutation
  size, which results in computation times an order of magnitude shorter than
  CEGO, which makes it specially useful when both computation time and number
  of evaluations are limited.  \sloppy
\end{abstract}
%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002950.10003624.10003625.10003630</concept_id>
<concept_desc>Mathematics of computing~Combinatorial optimization</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010061.10011795</concept_id>
<concept_desc>Theory of computation~Random search heuristics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}.
% , such as Kendall's-$\tau$ distance for permutations
% (Eq.~\ref{eq:kendall_decomp}).

In CEGO, the surrogate model is
explored by a GA with crossover and mutation operators appropriate for the
particular combinatorial problem. A thorough comparison of CEGO
with other optimizers for expensive black-box combinatorial optimization ranked
it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.

>>>>

% \author{Ekhiñe Irurozki\inst{1} \and Manuel López-Ibáñez\inst{2}\orcidID{0000-0001-9974-1295}}
% \institute{
%    Basque Center for Applied Mathematics\\
%    \email{eirurozki@bcamath.org}
%    \and
%    University of Málaga, Málaga, Spain\\
%    \email{manuel.lopez-ibanez@uma.es}
%  }
%    \renewcommand{\shortauthors}{B. Trovato et al.}
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Expensive black-box combinatorial optimization problems arise in practice
  when the objective function is evaluated by means of a simulator or a
  real-world experiment. Since each fitness evaluation is expensive in terms of
  time or resources, the number of possible evaluations is typically several
  orders of magnitude smaller than in non-expensive problems. Classical
  optimization methods are not useful in this scenario.
  In this paper, we propose and analyze UMM, an estimation-of-distribution
  (EDA) algorithm based on a Mallows probabilistic model and unbalanced rank
  aggregation (uBorda).
  Experimental results on black-box versions of LOP and PFSP show that UMM
  outperforms the solutions obtained by CEGO, a Bayesian optimization algorithm
  for combinatorial optimization. Nevertheless, a slight modification to CEGO,
  based on the different interpretations for rankings and orderings,
  significantly improves its performance, thus producing solutions that are
  slightly better than those of UMM and dramatically better than the original
  version.  Another benefit of UMM is that its computational complexity
  increases linearly with both the number of function evaluations and the
  permutation size, which results in computation times an order of magnitude
  shorter than CEGO, making it specially useful when both computation time and
  number of evaluations are limited.
  \sloppy
\end{abstract}
%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002950.10003624.10003625.10003630</concept_id>
<concept_desc>Mathematics of computing~Combinatorial optimization</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10010061.10011795</concept_id>
<concept_desc>Theory of computation~Random search heuristics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

CEGO replaces the Euclidean distance measure, used by the surrogate model in
EGO, with a distance measure appropriate to combinatorial
landscapes~\citep{ZaeStoBar2014:ppsn}.
% , such as Kendall's-$\tau$ distance for permutations
% (Eq.~\ref{eq:kendall_decomp}).
%
In CEGO, the surrogate model is
explored by a GA with crossover and mutation operators appropriate for the
particular combinatorial problem. A thorough comparison of CEGO
with other optimizers for expensive black-box combinatorial optimization ranked
it as the best-performing~\citep{ZaeStoFriFisNauBar2014}.
